---
title: "W8_WebScraping"
author: "Nicole Dwenger"
date: "2020/11/01"
output: html_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, rvest, magrittr, scales, knitr, lubridate, tibble, gdata)
```

# Week 8: Webscraping 
I decided to scrape some data from [Spotify Charts](https://spotifycharts.com/). For this, I used this [Tutorial](https://datacritics.com/2018/03/20/scrape-it-yourself-spotify-charts/). 


## 1. Defining urls to scrape from
```{r eval = FALSE}

# defining the base url
dk <- "https://spotifycharts.com/regional/dk/weekly/" # denmark
de <- "https://spotifycharts.com/regional/de/weekly/" # germany
us <- "https://spotifycharts.com/regional/us/weekly/" # us
uk <- "https://spotifycharts.com/regional/gb/weekly/" # uk

# defining the variation in the ulr (wich is the week)
timevalues1 <- seq(as.Date("2019/12/27"), as.Date("2020/10/23"), by = "week")
timevalues2 <- seq(as.Date("2020-01-03"), as.Date("2020/10/30"), by = "week")

# to make the final urls from which I will be scaping, I add together the base + variation 
unitedata <- function(country, timevalue1, timevalue2){
 full_url <- paste0(country, timevalue1, "--", timevalue2)
 full_url
}

# for each country I apply the function to get all the final urls I want to scrape from
fullurl_dk <- unitedata(dk, timevalues1, timevalues2)
fullurl_de <- unitedata(de, timevalues1, timevalues2)
fullurl_us <- unitedata(us, timevalues1, timevalues2)
fullurl_uk <- unitedata(uk, timevalues1, timevalues2)

```

## 2. Defining function to use for scraping and applying it 
```{r eval = FALSE}

spotify_scrape <- function(x){
 page <- x
 rank <- page %>% read_html() %>% html_nodes('.chart-table-position') %>% html_text() %>% as.data.frame()
 track <- page %>% read_html() %>% html_nodes('strong') %>% html_text() %>% as.data.frame()
 artist <- page %>% read_html() %>% html_nodes('.chart-table-track span') %>% html_text() %>% as.data.frame()
 streams <- page %>% read_html() %>% html_nodes('td.chart-table-streams') %>% html_text() %>% as.data.frame()
 date <- page %>% read_html() %>% html_nodes('.responsive-select-value') %>% html_text() %>% .[3] 
 url <- page %>% read_html() %>% html_nodes('td.chart-table-image a') %>% html_attr('href')
 # combine, name, and make it a tibble
 chart <- cbind(rank, track, artist, streams, url)
 chart$date <- date
 names(chart) <- c("rank", "track", "artist", "streams", "url", "date")
 chart <- as_tibble(chart)
 return(chart)
}

# apply the function to scape the data
data_dk <- map_df(fullurl_dk, spotify_scrape)
data_de <- map_df(fullurl_de, spotify_scrape)
data_us <- map_df(fullurl_us, spotify_scrape)
data_uk <- map_df(fullurl_uk, spotify_scrape)
```

## 3. Cleaning data
```{r eval = FALSE}

# combine data from all countries, 
data <- combine(data_dk, data_de, data_us, data_uk) %>%
  # clean the columns and make some new ones from the existing ones
  mutate(country = str_extract(source, ".{2}$"),
         streams = as.numeric(gsub(",", "", streams)),
         artist = str_remove(artist, "by "),
         date = as.Date(date, "%m/%d/%Y"),
         month = str_extract(date, "\\b\\d{2}\\b"),
         track_id = str_remove(url, "https://open.spotify.com/track/")) %>%
  # and put it into a nice order
  select(rank,  
         track,
         artist,
         streams,
         date,
         month,
         country,
         url,
         track_id)

# calculate the total streams for each country and add to dataframe
total_country_streams <- data %>%
  group_by(country, date) %>%
  summarise(total_country = sum(streams), .groups = "drop_last")
data <- merge(data, total_country_streams, by = c("country", "date"))

# save data
write.csv(data, "W8_spotify_data.csv")

```

## 4. Top artists in each country
I am aware that here it would make sense to also take into account how many songs each artist had, but I ignored that here and just went for the streams of each artist. 
```{r}

# load data
data <- read_csv("W8_spotify_data.csv")

# disable scientific notation
options(scipen=999)

# make function for plot
top_artist_plot <- function(short_country) {
  data %>% 
     filter(country == short_country) %>%
     group_by(artist) %>% 
     summarise(total_artist = sum(streams), .groups = "drop_last") %>% 
     arrange(desc(total_artist)) %>%
     top_n(25, total_artist) %>%
 ggplot() +
     geom_col(aes(reorder(artist, total_artist), y = total_artist), fill = "lightblue3") +
     coord_flip() +
     labs(x = "Artist", y = "Total Streams", title = paste("Top Artists for Country:", short_country))
}

# apply for each country
for (country in unique(data$country)) {
  print(top_artist_plot(country))
}


```

## 5. Top songs in each country
I am aware that here it would make sense to also take into account how many songs each artist had, but I ignored that here and just went for the streams of each artist. 
```{r}

# disable scientific notation
options(scipen=999)

# make function for plot
top_track_plot <- function(short_country) {
  data %>% 
     filter(country == short_country) %>%
     group_by(track) %>% 
     summarise(total_track = sum(streams), .groups = "drop_last") %>% 
     arrange(desc(total_track)) %>%
     top_n(25, total_track) %>%
 ggplot() +
     geom_col(aes(reorder(track, total_track), y = total_track), fill = "lightcoral") +
     coord_flip() +
     labs(x = "Track", y = "Total Streams", title = paste("Top Tracks for Country:", short_country))
}

# apply for each country
for (country in unique(data$country)) {
  print(top_track_plot(country))
}


```

## 6. Development of song "Blinding Lights" 
I saw that the song was the Nr. 1 for 3 of the countries, and Nr. 2 for the us, so I decided to plot how the number of streams developed in the countries. 
```{r}
data %>%
  filter(track == "Blinding Lights") %>%
  ggplot() +
    geom_point(aes(y = streams, x = date, color = country)) +
    geom_line(aes(y = streams, x = date, color = country)) +
    scale_y_log10() +
    labs(x = "Time (in 2020)", y = "Total Streams", color = "Country", title = "Streams of 'Blindings Lights'")
    

data %>%
  filter(track == "Blinding Lights") %>%
  group_by(date) %>% 
  summarise(total = sum(streams), .groups = "drop_last") %>% 
  ggplot(aes(y = total, x = date)) +
    geom_point() +
    geom_line() +
    labs(x = "Time (in 2020)", y = "Total Streams", title = "Streams of 'Blindings Lights' across countries")


```

#7. Development of 5 songs across all countries
```{r}

data %>%
  filter(track == "Roses - Imanbek Remix" | track == "The Box" | track == "Blinding Lights" | track == "ROCKSTAR (feat. Roddy Ricch)" | track == "Dance Monkey") %>%
  group_by(track, date) %>% 
  summarise(total = sum(streams), .groups = "drop_last") %>%
  ggplot(aes(y = total, x = date, color = track, group = track)) +
    #geom_point() +
    geom_line() +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme(legend.position = c(0.8, 0.8)) +
    labs(x = "Time (in 2020)", y = "Total Streams", color = "Track", title = "Streams of 5 Top Tracks across countries")

```


