---
title: "HW6 - Loops + Functions"
author: "Nicole Dwenger"
date: "25/10/2020"
output: html_document
---

## Preparations

```{r libraries, message = FALSE}

# load packages
library(tidyverse)
library(gapminder)
library(stringr)

# save data as data frame
data <- as.data.frame(gapminder)

```

## Task 1
**Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.**

```{r}

# definsive function 
gdp_calc <- function(data, country, year){
  stopifnot(country %in% data$country, # make sure the country is in the data
            year %in% data$year[data$country == country], # year for the country is in the data
            is.numeric(data$gdpPercap)) # make sure the GDP is numeric
  subset <- data[data$country == country & data$year == year,] # make subset of country and yeear
  gdp <- subset$gdpPercap*subset$pop # calcualte gpd 
  return(paste("GDP of", country, "in", year, "was", gdp)) #return it in a nice sentence
}

# calculate GDP for denmark in the following years
gdp_calc(data, "Denmark", 1967)
gdp_calc(data, "Denmark", 1977)
gdp_calc(data, "Denmark", 1987)
gdp_calc(data, "Denmark", 2007)
# gdp_calc(data, "Denmark", 2017) there is no value for 2017 in the data for Denmark


```


## Task 2
**Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.**
*Note: I was unsure whether life expectancy here means "mean life expectancy" (just like in task 3) or whether it means "one of the life expectancy values" - but I assumed the first one.* 

```{r}

# print the life expectancy test if a country does start with B
for (country in unique(data$country)) {
  
  # if a country starts with B
  if (str_detect(country, "^B") == TRUE) {
    # calculate the mean life expectancy 
    mean_lifeExp <- mean(data$lifeExp[data$country == country])
    # and depending on the result print different things
    lifeExp <- ifelse(mean_lifeExp < 50, "smaller than 50.",
                  ifelse(mean_lifeExp > 50 & mean_lifeExp < 70, "between 50 and 70.", 
                  "greater than 70."))
    print(paste("In", country, "the mean life expectancy is", lifeExp))
    }
  }
 
```


## Optional: Task 3
**Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘M’ and graphs life expectancy against time (using plot() function) as a line graph if the mean life expectancy is under 50 years.**
```{r}

for (country in unique(data$country)) {
  if (str_detect(country, "^M") == TRUE & mean(data$lifeExp[data$country == country]) < 50) {
    subset <- data[data$country == country,]
    plot(y = subset$lifeExp, x = subset$year, 
         type = "l", 
         main = country, 
         ylab = "Life Expectancy in Years", 
         xlab = "Time (Year)")
  }
}

```



